\texttt{yacoub}: a Python package for Simulating Generalized Fading Channels

### 2018-03-15 paper/yacoub.tex
add more maths

\begin{abstract}
    We present a well tested Python-based library for simulating and computing
    generalized fading channels, named \texttt{yacoub}. We describe the
    applicability of \texttt{yacoub} using examples in recent communications
    systems challenges, namely: spectrum sensing, bit error rate computation,
    and parameter estimation in generalized fading channels. For the latter,
    a Majorization-Minimization algorithm is derived. The development of
    \texttt{yacoub} open source and its code is avaliable at
    \texttt{http://github.com/mirca/yacoub}.
\end{abstract}

@@

\subsection{Note on notation}
Scalars and random variables are denoted as \textit{italic} small-case letters \textit{e.g.} $x$;
vectors and random vectors are denoted as \textit{italic}, boldface, small-case letters \textit{e.g.} $\bm{x}$.
The $n-$th component of a vector $\bm{x}$ is denoted as $x_n$.
A complex vector of length $n$ is defined as $\bm{x} \in \mathbb{C}^{n\times 1}$. All
vectors are column vectors. Matrices are denoted as \textit{italic}, boldface, capital
letters as in $\bm{X}$; the identity matrix of order $n$ is denoted as $\bm{I}_n$.
We define a discrete-time circularly symmetric Gaussian process $\bm{z}$ as any (finite or infinite) collection
of random varibles $\bm{z}~=~\bm{x}~+~j\bm{y}$, $j \triangleq \sqrt{-1}$, such that
$\bm{x}$ and $\bm{y}$ are iid jointly Gaussian with zero mean vector and covariance
matrix given by $\mathbb{E}\left[\bm{z}\bm{z}^{\dagger}\right]$, in which $\bm{z}^\dagger$
means the conjugate transpose of $\bm{z}$. The expectation value wrt a random variable $x$
is denoted as $\mathbb{E}_x$. The probability of an event $A$ is denoted as $\mathbb{P}(A)$.
The indicator function is denoted as $\mathbb{I}(\cdot)$ and it evaluates to $1$ if its
argument is true and $0$ otherwise.

@@

@@

@@

Given a vector $\bm{h} \in \mathbb{R}^{n \times 1}$ of iid samples from the
$\alpha-\mu$ distribution, we would like to estimate the value of $\alpha$.
The pdf of $\bm{h}$ is given as
\begin{align}
    p(\bm{h}) = \prod_{i=1}^{n}p(h_i) = \prod_{i=1}^{n}\dfrac{\alpha \mu^\mu h_i^{\alpha\mu -1}}
    {\Gamma(\mu)}\exp\left(-\mu h_i^\alpha\right)\\
    = \left(\dfrac{\alpha \mu^\mu }{\Gamma(\mu)}\right)^{n}
    \exp\left(-\mu\sum_{i=1}^{n}h_i^{\alpha}\right)\prod_{i=1}^{n}h_i^{\alpha\mu -1}
\end{align}
and the log likelihood function (up to an additive constant that does not depend
upon $\alpha$) is given as
\begin{align}
    \log p(\bm{h}) \cong n\log\alpha - \mu\sum_{i=1}^{n}h_i^{\alpha} + \alpha\mu\sum_{i=1}^{n}\log h_i
\end{align}
Noting that
\begin{equation}
    \sum_{i=1}^{n}h_i^{\alpha} \geq \sum_{i=1}^{n} h_i^{\alpha_t} + (\alpha - \alpha_t)\sum_{i=1}^{n}h_i^{\alpha_t}\log h_i
\end{equation}
and therefore
\begin{align}
    \log p(\bm{h}) \leq n\log\alpha - \mu\left(\sum_{i=1}^{n} h_i^{\alpha_t} +
    (\alpha - \alpha_t)\sum_{i=1}^{n}h_i^{\alpha_t}\log h_i\right) + \alpha\mu\sum_{i=1}^{n}\log h_i
\end{align}

>>>>

\begin{abstract}
    We present a well tested Python-based library for simulating and computing
    generalized fading channels, named \texttt{yacoub}. We describe the
    applicability of \texttt{yacoub} using examples in recent communications
    systems challenges, namely: spectrum sensing, bit error rate computation,
    and parameter estimation in generalized fading channels. The development of
    \texttt{yacoub} open source and its code is avaliable at
    \texttt{http://github.com/mirca/yacoub}.
\end{abstract}

@@

\subsection{Note on notation}
Scalars and random variables are denoted as \textit{italic} small-case letters \textit{e.g.} $x$;
vectors and random vectors are denoted as \textit{italic}, boldface, small-case letters \textit{e.g.} $\bm{x}$.
The $n-$th component of a vector $\bm{x}$ is denoted as $x_n$.
A complex vector of length $n$ is defined as $\bm{x} \in \mathbb{C}^{n\times 1}$. All
vectors are column vectors. Matrices are denoted as \textit{italic}, boldface, capital
letters as in $\bm{X}$; the identity matrix of order $n$ is denoted as $\bm{I}_n$.
We define a discrete-time circularly symmetric Gaussian process $\bm{z}$ as any collection
of random varibles $\bm{z}~=~\bm{x}~+~j\bm{y}$, $j \triangleq \sqrt{-1}$, such that
$\bm{x}$ and $\bm{y}$ are iid jointly Gaussian with zero mean vector and covariance
matrix given by $\mathbb{E}\left[\bm{z}\bm{z}^{\dagger}\right]$, in which $\bm{z}^\dagger$
means the conjugate transpose of $\bm{z}$. The expectation value wrt a random variable $x$
is denoted as $\mathbb{E}_x$. The probability of an event $A$ is denoted as $\mathbb{P}(A)$.
The indicator function is denoted as $\mathbb{I}(\cdot)$ and it evaluates to $1$ if its
argument is true and $0$ otherwise.

@@

\section{Majorization-Minimization Algorithms}

@@

\subsection{Parameter Estimation in Nakagami-$m$ fading}
\begin{align}
p(\bm{h}) = & \prod_{i=1}^{n}\dfrac{2m^m}{\Gamma(m)\Omega^{m}}h_i^{2m - 1}
              \exp\left(-\dfrac{mh_i^2}{\Omega}\right) \\
          = & \left(\dfrac{2m^m}{\Gamma(m)\Omega^{m}}\right)^{n}
          \exp\left(-\dfrac{m\sum_{i=1}^{n}h_i^2}{\Omega}\right) \prod_{i=1}^{n}h_i^{2m - 1}
\end{align}
\begin{align}
    \log p(\bm{h}) \cong &~n\left(m\left(\log m - \log\Omega\right) - \log\Gamma(m)\right)\nonumber
    \\ & -m\left(\dfrac{\sum_{i=1}^{n}h_i^2}{\Omega} - 2\sum_{i=1}^{n}\log h_i\right)
    \label{eq:loglike}
\end{align}

A direct maximum likelihood estimator for (\ref{eq:loglike}) has been investigated to be infeasible~\cite{paper}.

Therefore, we use a Majorization-Minimization algorithm to find smooth
and easy to optimize upper bounds for $\log p(\bm{h})$.

Note that,
\begin{align}
    -\log\Gamma(m) \geq &-\log\Gamma(m_t) + (m - m_t) \nonumber \\
                        &+ \left(m_t - \dfrac{1}{2}\right)\log (m_t-1)\nonumber\\
                        &-\left(m - \dfrac{1}{2}\right)\log (m-1), m \geq m_t > 1
\end{align}
and
\begin{align}
    &m \log m \geq m\log m_t + m - m_t,~m > 0\\
    & - m\log (m-1) \geq - \log(m-1) - (m-1)\left(\log (m_t-1) + \dfrac{m-1}{m_t-1} - 1\right),~m > 1
\end{align}
(in all inequalities, equality is achieved iff $m = m_t$).

Therefore,
\begin{figure*}[!htb]
\begin{align}
    - \log p(\bm{h}) \leq -n&\left(m\log m_t + 2(m - m_t) - m\log\Omega - \log\Gamma(m_t) +
        \left(m_t - \dfrac{1}{2}\right)\log (m_t-1) - \dfrac{1}{2}\log(m-1)\right. \nonumber\\
        &\left.-(m-1)\left(\log (m_t-1) + \dfrac{m-1}{m_t-1} - 1\right)
        \right) +m\left(\dfrac{\sum_{i=1}^{n}h_i^2}{\Omega} - 2\sum_{i=1}^{n}\log h_i\right), m \geq m_t > 1.
\end{align}
\end{figure*}

\begin{figure*}[!htb]
    \begin{align}
    \dfrac{\partial g(m | m_t)}{\partial m} = -n\left(\log m_t + 2 - \log\Omega - \dfrac{1}{2(m - 1)}
    - \left(\log (m_t - 1) - 1\right) + \dfrac{-2m + 1}{m_t - 1} \right)
    + \dfrac{\sum_{i=1}^{n}h_i^2}{\Omega} - 2\sum_{i=1}^{n}\log h_i = 0
    \end{align}
\end{figure*}

\begin{figure*}[!htb]
    \begin{align}
        2(m - 1) \left(-n\left(\log m_t + 2 - \log\Omega - (\log(m_t - 1) - 1)\right) + \dfrac{\sum_{i=1}^{n}h_i^2}{\Omega} - 2\sum_{i=1}^{n}\log h_i \right) + n - \dfrac{2n}{m_t-1}(-2m^2 + 3m - 1) = 0\\
        2(m - 1) a + n - \dfrac{2n}{m_t-1}(-2m^2 + 3m - 1) = 0\\
        m^2\left(\dfrac{4n}{m_t - 1}\right) + m\left(\dfrac{6n}{m_t - 1} + 2a\right) -2a + n\left(\dfrac{m_t +1}{m_t - 1}\right) = 0\\
    \end{align}
\end{figure*}

@@

Given a vector $\bm{h} \in \mathbb{R}^{n \times 1}$ of iid samples from the
$\alpha-\mu$ distribution, we would like to estimate the value of...
The pdf of $\bm{h}$ is given as
\begin{align}
    p(\bm{h}) = \prod_{i=1}^{n}p(h_i) = \prod_{i=1}^{n}\dfrac{\alpha \mu^\mu h_i^{\alpha\mu -1}}
    {\Gamma(\mu)}\exp\left(-\mu h_i^\alpha\right)\\
    = \left(\dfrac{\alpha \mu^\mu }{\Gamma(\mu)}\right)^{n}
    \exp\left(-\mu\sum_{i=1}^{n}h_i^{\alpha}\right)\prod_{i=1}^{n}h_i^{\alpha\mu -1}
\end{align}
and the loglikelihood is given as
\begin{align}
    \log p(\bm{h}) \cong
    &~n\left(\log\alpha + \mu \log \mu - \log \Gamma(\mu)\right) \nonumber\\
    & - \mu\sum_{i=1}^{n}h_i^{\alpha} + \alpha\mu\sum_{i=1}^{n}\log h_i
\end{align}

Therefore, we want to solve
\begin{align}
    & \text{minimize} -\log p(\bm{h})\\
    & \text{s.t.}~\alpha > 0, \bm{h} > 0 \nonumber
\end{align}
